Data Migration under a Changing Schema in Ampersand

### 2023-10-18 2022Migration/articleMigrationFACS.tex
correct a mistake

\section{Generating a Migration Script}
   A migration script takes care of two things: copying data and ensuring that the migration system is a proper information system.
\subsection{Generating a migration script}
   In the migration system, we need to refer to the items (concepts, relations, and constraints) of both the existing system and the desired system.
   We have to relabel items with prefixes to avoid name clashes in the migration system.
   We use a left arrow to denote relabeling by prefixing the name of the item with ``old.''.
\begin{equation}
   \begin{array}[m]{rcl}
      \overleftarrow{\pair{\dataset}{\schema}}&=&\pair{\overleftarrow{\dataset}}{\overleftarrow{\schema}}\\
      \overleftarrow{\pair{\triples}{\instance}}&=&\pair{\overleftarrow{\triples}}{\instance}\\
      \overleftarrow{\quintuple{\concepts}{\rels}{\rules}{\transactions}{\busConstraints}}&=&\quintuple{\concepts}{\overleftarrow{\rels}}{\overleftarrow{\rules}}{\overleftarrow{\transactions}}{\overleftarrow{\busConstraints}}\\
      \overleftarrow{\triples}&=&\{\triple{a}{\overleftarrow{r}}{b}\mid\triple{a}{r}{b}\in\triples\}\\
      \overleftarrow{\declare{n}{A}{B}}&=&\declare{old.n}{A}{B}\\
      \overleftarrow{X}&=&\{\overleftarrow{x}\mid x\in X\}\\
      \viol{\overleftarrow{u}}{\overleftarrow{\dataset}}&=&\viol{u}{\dataset}\\
      \sign{\overleftarrow{u}}&=&\sign{u}\\
      \enfRel{\overleftarrow{u}}&=&\overleftarrow{\enfRel{u}}
   \end{array}
\end{equation}
   The notation $\overrightarrow{\infsys}$ is defined similarly, using prefix ``new.'' instead of ``old.''.

@@

@@

      \widetilde{\declare{n}{A}{B}}={}&\declare{(\id{copied}.n)}{A}{B}\\
      \begin{split}
         \transactions_1={}&\{\widetilde{r}\mapsfrom \popF{\overrightarrow{r}\cap\overleftarrow{r}} \mid r \in \rels_\schema \cap \rels_{\schema'}\}\\
         &\cup\{\overrightarrow{r}\mapsfrom \popF{\overleftarrow{r} - \widetilde{r}} \mid r \in \rels_\schema \cap \rels_{\schema'}\}
      \end{split}

@@

@@

   For each new blocking invariant $u$, we generate a helper relation: ${\tt fixed}_u$, and a transaction to produce its population:

@@

      \rels_\theenumi={}&\{{\tt fixed}_u\mid u \in \rules_{\schema'}-\rules_{\schema}\}\\

@@

@@

      \schema_\migrsys=\quintuple{\concepts_\dataset\cup\concepts_{\dataset'}}{\overleftarrow\rels_{\schema}\cup\overrightarrow\rels_{\schema'}\cup\rels_3}{\rules_\text{block}}{\transactions_1\cup\transactions_2}{\busConstraints_\text{fix}\cup\busConstraints_{\schema'}}

>>>>

\section{Generating a Migration Script}
   A migration script takes care of two things: copying data and ensuring that the migration system is a proper information system.
\subsection{Algorithm for generating a migration script}
   In the migration system, we need to refer to the items (concepts, relations, and constraints) of both the existing system and the desired system.
   We have to relabel items with prefixes to avoid name clashes in the migration system.
   We use a left arrow to denote relabeling by prefixing the name of the item with ``old.''.
\begin{equation}
   \begin{array}[m]{rcl}
      \overleftarrow{\pair{\dataset}{\schema}}&=&\pair{\overleftarrow{\dataset}}{\overleftarrow{\schema}}\\
      \overleftarrow{\pair{\triples}{\instance}}&=&\pair{\overleftarrow{\triples}}{\instance}\\
      \overleftarrow{\quintuple{\concepts}{\rels}{\rules}{\transactions}{\busConstraints}}&=&\quintuple{\concepts}{\overleftarrow{\rels}}{\overleftarrow{\rules}}{\overleftarrow{\transactions}}{\overleftarrow{\busConstraints}}\\
      \overleftarrow{\triples}&=&\{\triple{a}{\overleftarrow{r}}{b}\mid\triple{a}{r}{b}\in\triples\}\\
      \overleftarrow{\declare{n}{A}{B}}&=&\declare{old.n}{A}{B}\\
      \overleftarrow{X}&=&\{\overleftarrow{x}\mid x\in X\}\\
      \viol{\overleftarrow{u}}{\overleftarrow{\dataset}}&=&\viol{u}{\dataset}\\
      \sign{\overleftarrow{u}}&=&\sign{u}\\
      \enfRel{\overleftarrow{u}}&=&\overleftarrow{\enfRel{u}}
   \end{array}
\end{equation}
   The notation $\overrightarrow{\infsys}$ is defined similarly, using prefix ``new.'' instead of ``old.''.

@@

For every relation $r$ shared by the existing and desired systems, we generate a helper relation: ${\tt copy}_r$, and a transaction to produce its population.
We also generate a transaction to populate the relation $\overrightarrow{r}$ (in the desired system).

@@

      \rels_1={}&\{{\tt copy}_r\mid r \in \rels_{\schema'}\cap\rels_{\schema}\}\\
      \transactions_1={}&\{{\tt copy}_r\mapsfrom \popF{\overrightarrow{r}\cap\overleftarrow{r}} \mid r \in\rels_1\}\cup\{\overrightarrow{r}\mapsfrom \popF{\overleftarrow{r}-{\tt copy}_r} \mid r \in\rels_1\}

@@

   The helper relation is needed to ensure that the copying process terminates.

@@

   For each new blocking invariant $u$, we generate a helper relation: ${\tt fixed}_u$, to register all violations that are fixed.
   We also need a transaction to produce its population:

@@

      \rels_2={}&\{{\tt fixed}_u\mid u \in \rules_{\schema'}-\rules_{\schema}\}\\

@@

   Note that the blocking invariants from $\schema'\cap\schema$ are also used in $\migrsys$ to ensure continuity.

@@

      \schema_\migrsys=\langle{}&\concepts_\dataset\cup\concepts_{\dataset'},\\
      &\overleftarrow{\rels_{\schema}}\cup\overrightarrow{\rels_{\schema'}}\cup\rels_1\cup\rels_2,\notag\\
      &\rules_\text{block}\cup(\rules_{\schema}\cap\rules_{\schema'}),\notag\\
      &\transactions_1\cup\transactions_2,\notag\\
      &\busConstraints_\text{fix}\cup\busConstraints_{\schema'}\rangle\notag
